Nanocall: An Open Source Basecaller for Oxford Nanopore Sequencing Data

### 2016-03-17 paper-biorxiv/paper.tex
state transitions

\paragraph{State Transitions.}
The state transitions are prior probabilities of moving from one state to another state. The default state transitions are computed based on 2 parameters: the ``stay'' probability \pstay and the ``skip'' probability \pskip. The former, \pstay, is the probability that 2 consecutive events are emitted from the same context/state. So, this corresponds to a segmentation error where an erroneous event break was introduced. The latter, \pskip, is the probability that 2 consecutive events are emitted from 2 states that differ by more than 1 kmer shift. So, this corresponds to either a segmentation error or a sequencing error (i.e., DNA moved too fast) where 1 or more events were lost. A slight complication is that, by increasing the number of skips, there is always more than 1 way of going from any one state to any other. For example, \texttt{ACGTGT} can be followed by \texttt{GTGTAC} using either 1 or 3 skips. The exact formula for computing state transitions takes this into account this fact:

@@

By default, Nanocall uses $\pstay = .1$ and $\pskip = .3$, which was observed to work well on several data sets. When transition parameter training is enabled (see Supplement), the values of \pstay and \pskip are updated individually for each strand of each read, using the standard Baum-Welch algorithm.

>>>>

\paragraph{State Transitions.}

The state transitions are prior probabilities of moving from one state to another state. The default state transitions are computed based on two parameters: the ``stay'' probability \pstay and the ``skip'' probability \pskip (JTS note: typesetting error on pstay). The former, \pstay, is the probability that two consecutive events are emitted from the same context/state. This corresponds to a segmentation error where an erroneous event break was introduced. The latter, \pskip, is the probability that two consecutive events are emitted from states that differ by more than one kmer shift. This corresponds to either a segmentation error or a sequencing error (i.e., the DNA moved too quickly through the pore to register a detectable event) where one or more events were lost. A slight complication is that, by increasing the number of skips, there is always more than one way of going from any one state to any other. For example, \texttt{ACGTGT} can be followed by \texttt{GTGTAC} using either one or three skips. Our computation of the state transitions takes this into account:

@@

By default, Nanocall uses $\pstay = .1$ and $\pskip = .3$, which we observed to work well on several data sets. When transition parameter training is enabled (see Supplement), the values of \pstay and \pskip are updated individually for each strand of each read, using the Baum-Welch algorithm.
